---
title: "Personal Movement Analysis"
author: "Jim Perkins"
date: "November 13, 2016"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
---

```{r 1-setup_report, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)

#/ read the dataset
dataset<- read_csv("./activity.csv")
steps_by_day <- count(dataset, date, wt = steps)
```
#### Part 1 

> *For this part of the analysis, we ignored missing values (i.e. we stripped out any "NA" values*  
> *from the dataset before calculating the mean, median and the frequency plot.*

```{r 2-ignoreNA, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

steps <- as.data.frame(  steps_by_day[,2] )  
non.zero.steps <- steps$n[steps$n != 0]      #/remove all NA's from the data set 

x <- mean(non.zero.steps)
f.mean_steps_perday <- format(round(x,1),big.mark = ",")

y <- as.numeric(median(non.zero.steps))
f.median_steps_perday <- format(y, big.mark = ",")

```


Over the 2 month measurement period: AVERAGE steps per day was: **`r f.mean_steps_perday`** and the MEDIAN  steps per day was: **`r f.median_steps_perday`** 



Here is a frequency plot (i.e. histogram) of the total steps taken each day

```{r 3-print_Steps_Table, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
#/Histogram of steps
myTitle <- "Frequency plot of daily steps"
xlabel <- "Daily steps (in intervalS of 2,000)"
hist(non.zero.steps, breaks = 10, col = "light blue", xlab = xlabel, main = myTitle)

```

> *Meaning: The most frequently occurring steps in a day are between 10,000 to 12,000*  


```{r 4-plot_aveSteps_perInterval, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

x <- select(dataset, interval, date, steps) %>% filter(steps >0) 
tmp <- arrange(x, interval)
plotdata <- aggregate(x = tmp, by = list(tmp$interval), FUN = "mean")
title <- "Time series of ave. steps by interval (across all intervals)"
plot(plotdata$interval, plotdata$steps, type="l", main=title, col="dark green" ) 

```

 Which interval and day has the highest average steps?
```{r 5-highest.Interval, include=TRUE}
knitr::opts_chunk$set(echo =TRUE)
plotdata[which.max(plotdata$steps),]

```


 
 

 